### 2014-05-09 ApTranscriptome_TR.Rmd
Updated analysis of temperature gene activation for *Low* genes

Predict for responsive transcripts

```{r responsive_pred}

predFunc <- function(foo) {
  # Function supplied to `ddply` to predict expression at each temperature
  #
  # Args:
  #  foo: data.frame model is fit to
  #
  #  'model' must be specified in environment  
  #
  # Returns:
  #  data.frame of Transcripts with pval

  lmout <- eval(parse(text = paste("lm(", model, ", data = foo)", sep = "")))

  # predict values
  pout <- predict(lmout)
  # back-transform from log-scale
  foo$pTPM <- exp(pout)

  # prediction can give values very close to zero or negative. biologically, these are meaningless so change to zero
  foo$pTPM <- ifelse(foo$pTPM < 0, 0, round(foo$pTPM, 5))
  # plot
  #with(foo, qplot(val, TPM, colour = colony, geom = "smooth"))
  #with(foo, qplot(val, pTPM, colour = colony, geom = "smooth"))

  pred.out.df <- data.frame(Transcript = foo$Transcript, TPM = foo$TPM, val = foo$val, colony = foo$colony, pTPM = foo$pTPM)
}

# apply predFunc to all responsive transcripts
resp.TPM.dt.sub.pred <- ddply(resp.TPM.dt.sub, .(Transcript), .inform="TRUE", predFunc)

# setkey to Transcript and colony 
resp.TPM.dt.sub.pred <- data.table(resp.TPM.dt.sub.pred)
setkey(resp.TPM.dt.sub.pred, Transcript, colony)

# transcripts expressed at *High* temperatures in A22
A22.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(A22.high, "A22")]
str(A22.high.TPM.dt.sub)

@@

cor.test(Ar.high.T_on$T_on, Ar.high.T_on$pT_on)
cor.test(A22.high.T_on$T_on, A22.high.T_on$pT_on)

t.test(Ar.high.T_on$T_on, A22.high.T_on$T_on)
t.test(Ar.high.T_on$pT_on, A22.high.T_on$pT_on)
```

Genes with increased expression at *High* temperatures are turned on at higher temperatures in *ApVT* than *AcNC*.

@@

# plot
png("T_on_high.png")
T_on_plot <- ggplot(Ap.high.T_on, aes(x=T_on, fill=colony)) + geom_density(alpha=0.2, position="identity")
T_on_plot + scale_y_continuous(name="Density") +
  scale_x_continuous(name=expression("Temperature"))
dev.off()

# plot predicted
png("pT_on_high.png")
pT_on_plot <- ggplot(Ap.high.T_on, aes(x=pT_on, fill=colony)) + geom_density(alpha=0.2, position="identity")
pT_on_plot + scale_y_continuous(name="Density") +
  scale_x_continuous(name=expression("Temperature"))
dev.off()
```

>>>>

Predict expression for responsive transcripts.

```{r responsive_pred}
# apply predFunc to all responsive transcripts
resp.TPM.dt.sub.pred <- ddply(resp.TPM.dt.sub, .(Transcript), .inform="TRUE", predFunc)

# setkey to Transcript and colony 
resp.TPM.dt.sub.pred <- data.table(resp.TPM.dt.sub.pred)
setkey(resp.TPM.dt.sub.pred, Transcript, colony)
```

Calculate `T_on` for *High* genes in each colony.

```{r T_on_high}
# transcripts expressed at *High* temperatures in A22
A22.high.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(A22.high, "A22")]
str(A22.high.TPM.dt.sub)

@@

cor.test(Ar.high.T_on$T_on, Ar.high.T_on$pT_on)
cor.test(A22.high.T_on$T_on, A22.high.T_on$pT_on)
```

Determine if `T_on` is greater in *A22* or *Ar* for *High* genes.

```{r T_on_high_ttest}
t.test(Ar.high.T_on$T_on, A22.high.T_on$T_on)
t.test(Ar.high.T_on$pT_on, A22.high.T_on$pT_on)
```
Genes with increased expression at *High* temperatures are turned on at higher temperatures in *ApVT* than *AcNC*.

Plot `T_on` for *High* genes.

@@

# plot
T_on_plot <- ggplot(Ap.high.T_on, aes(x=T_on, fill=colony)) + geom_density(alpha=0.2, position="identity")
T_on_plot + scale_y_continuous(name="Density") +
  scale_x_continuous(name=expression("Temperature"))

# plot predicted
pT_on_plot <- ggplot(Ap.high.T_on, aes(x=pT_on, fill=colony)) + geom_density(alpha=0.2, position="identity")
pT_on_plot + scale_y_continuous(name="Density") +
  scale_x_continuous(name=expression("Temperature"))
```

Repeat analysis for *Low* genes.

```{r T_on_low}
# transcripts expressed at *Low* temperatures in A22
A22.low.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(A22.low, "A22")]
str(A22.low.TPM.dt.sub)

intersect(A22.low.TPM.dt.sub, A22.high.TPM.dt.sub)
# Good - no overlap

# make data.frame for results
A22.low.T_on <- data.frame(Transcript = unique(A22.low.TPM.dt.sub$Transcript), colony = rep(A22.low.TPM.dt.sub$colony, length = length(unique(A22.low.TPM.dt.sub$Transcript))), T_on = NA, pT_on = NA)

for(i in unique(A22.low.TPM.dt.sub$Transcript)) {
    subdf <- A22.low.TPM.dt.sub[i]
    subdf <- subdf[which(subdf$val < 21), ]
    T_on <- subdf[median(which(diff(subdf$TPM) == max(diff(subdf$TPM))))+1, val]
    pT_on <- subdf[median(which(diff(subdf$pTPM) == max(diff(subdf$pTPM))))+1, val]
    A22.low.T_on[which(A22.low.T_on$Transcript == i), "T_on"] <- T_on
    A22.low.T_on[which(A22.low.T_on$Transcript == i), "pT_on"] <- pT_on
}

# repeat for Ar
Ar.low.TPM.dt.sub <- resp.TPM.dt.sub.pred[J(Ar.low, "Ar")]
str(Ar.low.TPM.dt.sub)

intersect(A22.low.TPM.dt.sub, A22.high.TPM.dt.sub)
# Good - no overlap

Ar.low.T_on <- data.frame(Transcript = unique(Ar.low.TPM.dt.sub$Transcript), colony = rep(Ar.low.TPM.dt.sub$colony, length = length(unique(Ar.low.TPM.dt.sub$Transcript))), T_on = NA, pT_on = NA)

for(i in unique(Ar.low.TPM.dt.sub$Transcript)) {
    subdf <- Ar.low.TPM.dt.sub[i]
    subdf <- subdf[which(subdf$val < 21), ]
    T_on <- subdf[median(which(diff(subdf$TPM) == max(diff(subdf$TPM))))+1, val]
    pT_on <- subdf[median(which(diff(subdf$pTPM) == max(diff(subdf$pTPM))))+1, val]
    Ar.low.T_on[which(Ar.low.T_on$Transcript == i), "T_on"] <- T_on
    Ar.low.T_on[which(Ar.low.T_on$Transcript == i), "pT_on"] <- pT_on
}

t.test(Ar.low.T_on$T_on, A22.low.T_on$T_on)
t.test(Ar.low.T_on$pT_on, A22.low.T_on$pT_on)
```

Genes with increased expression at *Low* temperatures are turned on at lower temperatures in *AcNC* than *ApVT*.

Plot `T_on` for *Low* genes.

```{r plot_low_T_on, echo=FALSE, eval=TRUE}
# prep data for plotting
Ap.low.T_on <- rbind(A22.low.T_on, Ar.low.T_on)

# plot
T_on_plot_low <- ggplot(Ap.low.T_on, aes(x=T_on, fill=colony)) + geom_density(alpha=0.2, position="identity")
T_on_plot_low + scale_y_continuous(name="Density") +
  scale_x_continuous(name=expression("Temperature"))

# plot predicted
pT_on_plot_low <- ggplot(Ap.low.T_on, aes(x=pT_on, fill=colony)) + geom_density(alpha=0.2, position="identity")
pT_on_plot_low + scale_y_continuous(name="Density") +
  scale_x_continuous(name=expression("Temperature"))
```
